BSOS

### 2014-06-08 Paper.tex
more things

\part{Booting Up-William}
BSOS, as a portable and abstracted system, must be capable of bootstrapping on a variety of diverse and divergent architectures. Although hundreds exist, ranging from giants like x86, POWER, and ARM to lesser known embedded architectures like AVR and MIPS, the vast majority of fall into one of two families: CISC and RISC.

CISC, or Complex Instruction Set Computing, began as a side effect of early computer development. As the (comparatively primitive) computers and calculators of the 1950s and 1960s became increasingly complex and multifunctional, they required new paradigms and methods of automation. Because languages like FORTRAN and ALGOL required a certain amount of boilerplate in order to operate, early system architects build complex control structures into the instruction sets of their machines, bridging the semantic gap between microcode-assembly and full-features programming languages.

>>>>

\part{Booting Up-William}
BSOS, as a portable and abstracted system, must be capable of bootstrapping on a variety of diverse and divergent architectures. Although hundreds exist, ranging from giants like x86, POWER, and ARM to lesser known embedded architectures like AVR and MIPS, the vast majority fall into one of two families: CISC and RISC.

CISC, or Complex Instruction Set Computing, began as a side effect of early computer development. As the (comparatively primitive) computers and calculators of the 1950s and 1960s became increasingly complex and multifunctional, they required new paradigms and methods of automation. Because languages like FORTRAN and ALGOL required a certain amount of boilerplate in order to operate, early system architects build complex control structures into the instruction sets of their machines, bridging the semantic gap between microcode-assembly and full-features programming languages. Because this approach to abstraction is relatively simple to implement, CISC quickly became the go-to approach for architecture design, eventually resulting in seminal processors like the MOS 6502, Zilog z80, and Intel 8086/8088.

Although CISC architectures are simple to implement, they have a number of downsides. Nowhere is this more apparent than in x86 family. AMD64, the most recent iteration in a long line of x86-based instruction sets, contains a truly massive number of opcodes, registers, and control structures. While this makes it very powerful and often very convenient, it also poses a major problem for programming and efficiency. Programmers who wish to write assembly for modern x86 platforms must memorize and utilize a vast array of confusing mnemonics, many of which are sparsely documented. Even compilers, which can take advantage of modern instructions and features through sheer force, have trouble fully exploiting x86's full feature set. Consequently, although programs can be specialized or drastically simplified through advanced structures within the architecture, they often perform poorly or inefficiently on CISC systems.

RISC (Reduced Instruction Set Computing), which began formal development in the 1980s, is a solution to many of CISC's problems. Instead of providing a large collection of instructions for programmers to utilize, RISC systems strip the instruction set down to a bare minimum of operation, leaving only a few dozen simple opcodes (``primitives"). Additionally, those opcodes that remain follow strict cycle restrictions, rarely taking more than a single DMC (data memory cycle) to complete. This has two major side effects: simpler assembly programs, and the necessity of more abstraction between the user environment and the bare hardware. Although RISC CPU development is more recent than comparable CISC development, it is by no means unimportant. Architectures like DEC Alpha, SPARC, POWER, and ARM have become adopted in wide range of capacities, including supercomputers, mainframes, and mobile devices.

Unfortunately, RISC is not without its downsides as well.
